---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
# **Reproducible Research: Course Project 1**  
### It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the “quantified self” movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

## The aims of this project:
+ Loading and preprocessing the data
+ What is mean total number of steps taken per day?
+ What is the average daily activity pattern?
+ Imputing missing values
+ Are there differences in activity patterns between weekdays and weekends?

## **Loading and preprocessing the data**
#### Loading and unzipping the data. Looking at the summary of the data set. 
```{r echo=TRUE} 
unzip("./activity.zip")
activityData <- read.csv("./activity.csv")
summary(activityData)
```

## **What is mean total number of steps taken per day?**
#### Counting the steps per day and assigning the result to the variable S. Plotting a histogram on its basis.
```{r echo=TRUE} 
S <- aggregate(steps ~ date, activityData, sum, na.rm=TRUE)
hist(S$steps)
```

#### Calculating the mean of steps per day. Calculating the median of steps per day.
```{r echo=TRUE} 
meanS <- mean(S$steps)
meanS
```

```{r echo=TRUE} 
medianS <- median(S$steps)
medianS
```

## **What is the average daily activity pattern?**
#### Creating a plot of of the 5-minute interval and the average number of taken steps.
```{r echo=TRUE} 
steps1 <- aggregate(steps~interval, data=activityData, mean, na.rm=TRUE)
plot(steps~interval, data=steps1, type="l")
```
#### Calculating which interval contains the maximum number of steps and assigning it to variable MAX.
```{r echo=TRUE} 
MAX <- stepsPerInterval[which.max(stepsPerInterval$steps),]$interval
MAX
```

## **Imputing missing values**



## **Are there differences in activity patterns between weekdays and weekends?**
