---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
# **Reproducible Research: Course Project 1**  
### It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the “quantified self” movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.
*Figures can be found in the figures folder*.

## The aims of this project:
+ Loading and preprocessing the data
+ What is mean total number of steps taken per day?
+ What is the average daily activity pattern?
+ Imputing missing values
+ Are there differences in activity patterns between weekdays and weekends?

## **Loading and preprocessing the data**
*Load the data (i.e. read.csv())*  
*Process/transform the data (if necessary) into a format suitable for your analysis*  
#### Loading and unzipping the data. Looking at the summary of the data set. 
```{r echo=TRUE} 
unzip("./activity.zip")
activityData <- read.csv("./activity.csv")
summary(activityData)
```

## **What is mean total number of steps taken per day?**
*Calculate the total number of steps taken per day*
*If you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram of the total number of steps taken each day*
*Calculate and report the mean and median of the total number of steps taken per day*
#### Counting the steps per day and assigning the result to the variable S. Plotting a histogram on its basis.
```{r echo=TRUE} 
S <- aggregate(steps ~ date, activityData, sum, na.rm=TRUE)
hist(S$steps)
```

#### Calculating the mean of steps per day. Calculating the median of steps per day.
```{r echo=TRUE} 
meanS <- mean(S$steps)
meanS
```

```{r echo=TRUE} 
medianS <- median(S$steps)
medianS
```

## **What is the average daily activity pattern?**
#### Creating a plot of of the 5-minute interval and the average number of taken steps.
```{r echo=TRUE} 
steps1 <- aggregate(steps~interval, data=activityData, mean, na.rm=TRUE)
plot(steps~interval, data=steps1, type="l")
```
#### Calculating which interval contains the maximum number of steps and assigning it to variable MAX.
```{r echo=TRUE} 
MAX <- stepsPerInterval[which.max(stepsPerInterval$steps),]$interval
MAX
```

## **Imputing missing values**
#### Calculating the total number of the missing values.
```{r echo=TRUE}
totalValuesMissings <- sum(is.na(activityData$steps))
totalValuesMissings
```
####  Creating the average number of steps per weekday and interval and assigning it to the variable av.
```{r echo=TRUE}
av <- ddply(clean, .(interval, day), summarize, Avg = mean(steps))
```

#### Create a dataset with the missing values and merging it with the average weekday interval for substitution.
```{r echo=TRUE}
nadata<- activity[is.na(activity$steps),]
newdata<-merge(nadata, av, by=c("interval", "day"))
```

#### Cleaning and merging the data set. Comparing the data.
```{r echo=TRUE}
newdata2<- newdata[,c(6,4,1,2,5)]
colnames(newdata2)<- c("steps", "date", "interval", "day", "DateTime")
mergeData <- rbind(clean, newdata2)
sumTable2 <- aggregate(mergeData$steps ~ mergeData$date, FUN=sum, )
colnames(sumTable2)<- c("Date", "Steps")
as.integer(mean(sumTable2$Steps))
```
#### Creating the histogram.
```{r echo=TRUE}
hist(sumTable2$Steps, breaks=5, xlab="Steps", main = "Total Steps per Day with NAs Fixed", col="Black")
hist(sumTable$Steps, breaks=5, xlab="Steps", main = "Total Steps per Day with NAs Fixed", col="Grey", add=T)
legend("topright", c("Imputed Data", "Non-NA Data"), fill=c("black", "grey") )
```

## **Are there differences in activity patterns between weekdays and weekends?**
#### Creating the datasets for either weekend day or weekday.
```{r echo=TRUE}
activityDataNoNA$date <- as.Date(strptime(activityDataNoNA$date, format="%Y-%m-%d"))
activityDataNoNA$day <- weekdays(activityDataNoNA$date)
for (i in 1:nrow(activityDataNoNA)) {
    if (activityDataNoNA[i,]$day %in% c("Saturday","Sunday")) {
        activityDataNoNA[i,]$day<-"weekend"}
    else{activityDataNoNA[i,]$day<-"weekday"}
    }
stepsByDay <- aggregate(activityDataNoNA$steps ~ activityDataNoNA$interval + activityDataNoNA$day, activityDataNoNA, mean)
```
#### Plotting a graph.
```{r echo=TRUE}
names(stepsByDay) <- c("interval", "day", "steps")
library(lattice)
xyplot(steps ~ interval | day, stepsByDay, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")
```
